Stack Safety is a Security Property

### 2021-04-15 tex/main.tex
More work on abstract and intro

We propose a formal characterization
of stack safety, phrased using concepts from language-based security: a
combination of an integrity property (``the private
state in each caller's stack frame is held invariant by the callee'')
and a confidentiality property (``the callee's behavior is insensitive to the
caller's private state'')\bcp{and something about WBCF...}.
% These properties are sensible even in the presence of
% abnormal control flow.\apt{Is this important enough to mention here?}
% BCP: I think not in the abstract, but maybe in the intro.
% We further apply\apt{what does this mean? we didn't
%  invent it} a straightforward notion of
% ``well-bracketed control flow'' that restricts control flow to the
% typical pattern of matched calls and returns.
%
We use our characterization to validate the stack-safety ``micro-policies''
proposed by~\citet{DBLP:conf/sp/RoesslerD18}.
Specifically, we use property-based random testing to check that Roessler
and Dehon's ``eager'' micro-policy, which catches violations as early as
possible, enforces a simple ``stepwise'' variant of our properties (and
detects several broken variants), and that a slightly repaired version of
their more performant ``lazy'' micro-policy corresponds to weaker and more
extensional variants of the properties.
%
% \BCP{I still don't think the next sentence really nails it.  And (sigh) I am
%   still not really clear yet on whether our definition applies / might apply
%   to other enforcement mechanisms or not.  Saying that our attacker model is
%   stronger does not help---that should make it easier for us to describe
%   their protections, not harder.  So I don't understand what we're saying
%   here.}%
% %
% While we do not rule out other enforcement mechanisms, our
% definitions are strongly informed by micropolicy systems, especially
% our attacker model, which is stronger than other enforcement mechanisms
% can enable.

@@

%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
The most abstract characterization stack safety that we are aware of is in
recent work by \citet{Skorstengaard+19}, which defines stack safety as the conjunction of two
properties:
{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
Informally, LSE says that the contents of
a caller's stack frame are not read and written while its callees (and their callees, etc.) are
executing, and WBCF says that callees always
return to the point just after their call (if they return at all).
Skorstengaard et al.{} formalize these properties by defining an idealized
machine in which they hold {\em by construction};
they then show that an enforcement
mechanism (in their case, hardware capabilities) guarantees stack safety, by
proving that an ordinary machine with enforcement can simulate the idealized machine.
%an enforcement mechanism (in their case, hardware
%capabilities)
%can then be shown to guarantee stack safety by exhibiting a fully abstract
%embedding function
%from the idealized machine to an ordinary machine equipped with this
%enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
%  understand that sentence.}
%%
%% While this work represents an important step in the right direction, we believe
%% there is still room to craft an operational characterization
%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
%% In other words, we seek formal criteria that can be applied
%% to a mechanism for protecting stack-based low-level programs to judge
%% whether or not it actually ``guarantees stack safety.''
%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
%% confused by the references to these types of properties, yet to be introduced?}
%% We begin with a strong
%% local state encapsulation property that closely models the behavior of eager
%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
%% is too strong to characterize stack safety; reasonable enforcement mechanisms
%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
%% to support this technique and argue that the new property, defined relative to an
%% arbitrary notion of observation. This property is still strong enough to prevent
%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
%% other reasonable enforcement mechanisms can enforce it.

In the present work, we also distinguish between protecting stack data (like
LSE) and
control flow (like WBCF).  But our formulations of both properties are
somewhat different from those of Skorstengaard et al.  Rather than
defining stack safety implicitly,
as ``what a certain idealized machine does,'' we formalize explicit properties
based on the technical framework of language-based
security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
  survey?}\fi.
Our core properties are \emph{stepwise stack integrity} and
\emph{stepwise stack confidentiality}.  WBCF is a desireable property, but
we can express safety in an information-flow sense without reference to it,
even safety from attacks that hijack control flow.  \BCP{On consideration
  (especially in light of the weird ``C confuses A by returning directly to
  A instead of returning to B so that B can finish up some stuff before
  returning to A'' example\footnote{Actually, this example is not so weird.
  If C, the attacker, can return a {\em result} directly to A, who is
  expecting a result from B, then that is clearly bad!}), I
agree now that we should discuss WBCF.  It should be kept as distinct as
possible from the ``data safety'' part, since it is not really an
information flow property.  But our treatment of it is interestingly
different from StkTokens, because of our call annotations, and we should
discuss that.}

\bcp{This para could be clearer.  (I can try to fix it later but I want to
  keep reading now...)}In formalizing these security concepts, we treat {\em each} caller as needing
protection from its callee (which, in turn, needs protection from its
callee, etc.).
This is novel in that the
confidentiality property applies to many nested sub-traces within a single execution,
as opposed to traces of entire executions.
{\em Integrity} says that each caller's local data is protected from
modification by its callee(s),
motivated by the reasoning principle that local data persists across a call. Its
{\em stepwise} interpretation\bcp{We're introducing ``stepwise'' in a pretty
abrupt way---potentially confusing.} formalizes this very naturally: a caller's local data
is never modified during a call. {\em Confidentiality} says that a caller's
data is not leaked into its callee, and  that the callee cannot see uninitialized data.
As a reasoning principle, confidentiality guarantees that the callee's
behavior is determined only by a predictable interface (function arguments and
register contents) that does not include the caller's
local state or the uninitialized portion of the stack. This is formalized as a
{\em noninterference}-style property that, rather than comparing pairs of whole program traces,
compares a pair of sub-traces for each call, within other calls\bcp{?}. Once
again
{\em stepwise} describes the granularity of enforcement\bcp{?}: at every step of a call,
changes in the state are independent of the caller's data.

@@

@@

We lift the policy step function to operate on such pairs
by combining it with the regular step function for machine states.
%
\judgmenttwo{\(\mach \stepsto \mach'\)}{\((\mach, \pol)
               \polstep \pol'\)}
            {\(\mpstate{\mach}{\pol} \stepstopol
               \mpstate{\mach'}{\pol'}\)}
%
Note that the base machine has a total step function, and the policy-enhanced machine
refines it as a partial step. The policy state never changes behavior, but can cause
premature termination, which represents failstop behavior.

>>>>

We propose a formal characterization
of stack safety, formulated with concepts from language-based security: a
combination of an integrity property (``the private
state in each caller's stack frame is held invariant by the callee'')
a confidentiality property (``the callee's behavior is insensitive to the
caller's private state''), and a well-bracketedness property (``each
callee returns control to its immediate caller'').
% These properties are sensible even in the presence of
% abnormal control flow.\apt{Is this important enough to mention here?}
% BCP: I think not in the abstract, but maybe in the intro.
% We further apply\apt{what does this mean? we didn't
%  invent it} a straightforward notion of
% ``well-bracketed control flow'' that restricts control flow to the
% typical pattern of matched calls and returns.
%
We use these properties to validate the stack-safety ``micro-policies''
proposed by~\citet{DBLP:conf/sp/RoesslerD18}.  Specifically, we check (with
property-based random testing) that Roessler and Dehon's ``eager''
micro-policy, which catches violations as early as possible, enforces a
simple ``stepwise'' variant of our properties and correctly detects several
broken variants, and that (a repaired version of) their more performant
``lazy'' micro-policy corresponds to a slightly weaker and more extensional
variant of our properties.
%
% \BCP{I still don't think the next sentence really nails it.  And (sigh) I am
%   still not really clear yet on whether our definition applies / might apply
%   to other enforcement mechanisms or not.  Saying that our attacker model is
%   stronger does not help---that should make it easier for us to describe
%   their protections, not harder.  So I don't understand what we're saying
%   here.}%
% %
% While we do not rule out other enforcement mechanisms, our
% definitions are strongly informed by micropolicy systems, especially
% our attacker model, which is stronger than other enforcement mechanisms
% can enable.

@@

%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
The most abstract characterization stack safety that we are aware of is
recent work by \citet{Skorstengaard+19b,Skorstengaard+19}, which defines stack safety as the conjunction of two
properties:
{\em local state encapsulation (LSE)} and {\em well-bracketed control flow (WBCF)}.
Informally, LSE says that the contents of
a caller's stack frame are not read and written while its callees (and their callees, etc.) are
executing, and WBCF says that callees always
return to the point just after their call (if they return at all).
Skorstengaard et al.{} formalize these properties by defining an idealized
machine in which they hold {\em by construction};
they then show that an enforcement
mechanism (in their case, hardware capabilities) guarantees stack safety, by
proving that an ordinary machine with enforcement simulates the idealized
machine.
%an enforcement mechanism (in their case, hardware
%capabilities)
%can then be shown to guarantee stack safety by exhibiting a fully abstract
%embedding function
%from the idealized machine to an ordinary machine equipped with this
%enforcement mechanism.  \bcp{Many OOPSLA readers/reviewers will not
%  understand that sentence.}
%%
%% While this work represents an important step in the right direction, we believe
%% there is still room to craft an operational characterization
%% of stack safety. \apt{Perhaps our rhetoric should be more micro-policy-specific here?}
%% In other words, we seek formal criteria that can be applied
%% to a mechanism for protecting stack-based low-level programs to judge
%% whether or not it actually ``guarantees stack safety.''
%% \rb{The remainder of this paragraph may be diving too quickly, would a reader be
%% confused by the references to these types of properties, yet to be introduced?}
%% We begin with a strong
%% local state encapsulation property that closely models the behavior of eager
%% micro-policies from \citet{DBLP:conf/sp/RoesslerD18}. This ``eager'' property
%% is too strong to characterize stack safety; reasonable enforcement mechanisms
%% exist that violate it. One such mechanism is a ``lazy'' tag policy, similar to
%% the one in \citet{DBLP:conf/sp/RoesslerD18}. We weaken local state encapsulation
%% to support this technique and argue that the new property, defined relative to an
%% arbitrary notion of observation. This property is still strong enough to prevent
%% the bad behaviors we associate with unsafe stack behavior, but weak enough that
%% other reasonable enforcement mechanisms can enforce it.

In the present work, we also aim to protect both stack data (like
LSE) and
control flow (like WBCF).  But our formulations of these properties are
somewhat different from those of Skorstengaard et al.  Rather than
defining stack safety implicitly,
as ``what a certain idealized machine does,'' our properties are
based on the technical framework of language-based
security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
  survey?}\fi.
Our core properties are \emph{stepwise stack integrity},
\emph{stepwise stack confidentiality}, and \emph{return to caller}.
% \BCP{On consideration
%   (especially in light of the weird ``C confuses A by returning directly to
%   A instead of returning to B so that B can finish up some stuff before
%   returning to A'' example\footnote{Actually, this example is not so weird.
%   If C, the attacker, can return a {\em result} directly to A, who is
%   expecting a result from B, then that is clearly bad!}), I
% agree now that we should discuss WBCF.  It should be kept as distinct as
% possible from the ``data safety'' part, since it is not really an
% information flow property.  But our treatment of it is interestingly
% different from StkTokens, because of our call annotations, and we should
% discuss that.}

\bcp{This para could be clearer.  (I can try to fix it later but I want to
  keep reading now...)}In formalizing these security concepts, we treat {\em each} caller as needing
protection from its immediate callees (which, in turn, need protection from
their immediate callees, etc.).
This approach is novel in that the
confidentiality\bcp{only?} property applies to many nested sub-traces within a single execution,
as opposed to traces of entire executions.
{\em Integrity} says that each caller's local data is protected from
modification by its callee(s),
motivated by the reasoning principle that local data persists across a
call.

For the confidentiality and integrity properties, we consider two variants.
The
{\em stepwise} variants formalize the intuition that a caller's
local data
is never modified during a call. {\em Confidentiality} says that a caller's
data is not leaked into its callee, and  that the callee cannot see uninitialized data.
As a reasoning principle, confidentiality guarantees that the callee's
behavior is determined only by a predictable interface (function arguments and
register contents) that does not include the caller's
local state or the uninitialized portion of the stack. This is formalized as a
{\em noninterference}-style property that, rather than comparing pairs of whole program traces,
compares a pair of sub-traces for each call, within other calls\bcp{?}. Once
again
{\em stepwise} describes the granularity of enforcement\bcp{?}: at every step of a call,
changes in the state are independent of the caller's data.

@@

% BCP: This will intentionally fail when someone adds the real intended
% definition of \WF!
\newcommand{\WF}{\mathit{WF}}

@@

We lift the policy step function to operate on such pairs
by combining it with the regular step function for machine states.
%
\judgmenttwo{\(\mach \stepsto \mach'\)}{\((\mach, \pol)
               \polstep \pol'\)}
            {\(\mpstate{\mach}{\pol} \stepstopol
               \mpstate{\mach'}{\pol'}\)}
%
Note that the base machine has a total step function, while the policy-enhanced machine
refines it to a partial function. The extra policy state never changes behavior,
but it can cause
premature termination, representing failstop behavior.
